A Lattice Boltzmann Method for immiscible multiphase flow simulations using the Level Set Method

### 2015-08-19 Vortrag/folien.tex
Slides..

\visible<1->{
\begin{figure}[h!]
\includegraphics[width=4cm]{oilwater.jpg}
\end{figure}
}
\end{minipage}
&
\begin{minipage}{0.6\textwidth}
\visible<2->{
\begin{figure}[h!]
  \includegraphics[width=5.5cm]{bubbles.jpg}
\end{figure}
}
\end{minipage}
\end{tabular}

@@

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface capturing}
The interface between fluid phases is captured by a Level-Set Method. 
I.e. a \textit{level set function} $\varphi:= \varphi(x,t) \rightarrow \mathbb{R}$ is tracked through the fluid domain. The interface is given (implicitly) by the zero-isosurface of this function.
It holds:
$$ \frac{\partial \varphi}{\partial t} + \vec v \cdot \nabla \varphi = 0$$
where $\vec v$ is the velocity at the interface, given by NSE.

@@

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface implementation (Methods Coupling)}
Hydrodynamics are solved by LBM (D2Q9, SRT for our first test)
\begin{itemize}
\item<1-> Interface becomes a new boundary condition for LBM
$$f_i(x,t+1) = f_{i^*}^{+}(x,t) + 6 \Delta x f_{i}^{*}c_i \cdot \tilde{v} + R_i$$
\item<2-> $\tilde{v}$ is the velocity on the interface along the direction $c_i$
\end{itemize}
\begin{tabular}{l l}
\hspace{-.35cm} \begin{minipage}{0.6\textwidth}
\visible<1->{
\begin{itemize}
\item<3-> $R_i$ ensures the jump conditions of the normal stress and corrects the error terms resulting from the bounce back treatment
\end{itemize}
}
\end{minipage}
&
\begin{minipage}{0.4\textwidth}
\begin{figure}[h!]
\includegraphics[width=3cm]{skizze2.png}
\end{figure}
\end{minipage}
\end{tabular}
\end{frame}

@@

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Algorithm for LBM with level set}
\begin{itemize}
\item<1-> Create initial interface
\item<2-> Run level set method to calculate surface description
\item<3-> Run LBM for a prescribed number of steps
\item<4-> Run level set method for the same number of steps
\item<5-> Repeat till end of simulation
\end{itemize}
\end{frame}

\section{Results}
\begin{frame}
\frametitle{Validation}
TODO: Hier könnte man Bilder von unserer ersten Simulation zeigen. Leider können wir die Richtigkeit bisher nicht mit Zahlen belegen.
\framesubtitle{Validation setups}
\end{frame}

\section{Outlook}
\begin{frame}
\frametitle{Outlook}
$\rightarrow$ \textit{Conclusion habe ich rausgenommen. Ich wüsste zumindest nicht was wir da reinschreiben könnten. Wenn du was weißt, darfst du das aber sehr gerne wieder einfügen.}

\vspace{.8cm} 
Outlook:
\begin{itemize}
\item<1-> Add correction term to prevent mass loss
\item<2-> Reduce computational effort: Store Level-Set function only in narrow band around interface, Adalsteinsson and Sethian
TODO: Quellen als Footnotes + Uebersichtsfolie
\item<3-> Include thermal flow (simulate e.g. lava lamp) / Include gravity
\end{itemize}

@@

\begin{frame}
\frametitle{References}
\begin{thebibliography}{etw}
   \bibitem{thoemmes}
     Thömmes, Guido, et al. "A lattice Boltzmann method for immiscible multiphase flow simulations using the level set method." Journal of Computational Physics 228.4 (2009): 1139-1156
   \bibitem{mitchell}
     Mitchell, Ian, A Toolbox of Level Set Methods
\end{thebibliography}
\end{frame}

>>>>

\visible<1->{
\begin{figure}[h!]
\includegraphics[width=4cm]{oilwater.jpg}
\end{figure}
}
\end{minipage}
&
\begin{minipage}{0.6\textwidth}
\visible<2->{
\begin{figure}[h!]
  \includegraphics[width=5.5cm]{bubbles.jpg}
\end{figure}
}
\end{minipage}
\end{tabular}
\end{frame}

@@

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface capturing}
The interface between fluid phases is captured by a Level-Set Method. 
I.e. a \textit{level set function} $\varphi:= \varphi(x,t) \rightarrow \mathbb{R}$ is tracked through the fluid domain. The interface is given (implicitly) by the zero-isosurface of this function.
It holds:
$$ \frac{\partial \varphi}{\partial t} + \vec v \cdot \nabla \varphi = 0$$
where $\vec v$ is the velocity at the interface; given by NSE.

@@

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Interface implementation (Methods Coupling)}
Hydrodynamics are solved by LBM (Stream-Collide, D2Q9, SRT for our first test)
\begin{itemize}
\item<1-> Interface becomes a new boundary condition for LBM
  $$f_i(x,t+1) = f_{i^*}^{+}(x,t) + 6 \Delta x f_{i}^{*}c_i \cdot \tilde{v} + R_i$$
\end{itemize}
\begin{tabular}{l l}
\hspace{-.35cm} \begin{minipage}{0.65\textwidth}
\visible<1->{
\begin{itemize}
\item<2-> $\tilde{v}$ is the velocity on the interface along the direction $c_i$
\item<3-> $R_i$ ensures the jump conditions of the normal stress and corrects the error terms resulting from the bounce back treatment
\end{itemize}
}
\end{minipage}
&
\hspace{-.5cm} \begin{minipage}{0.4\textwidth}
\begin{figure}[h!]
\includegraphics[width=3cm]{skizze2.png}
\end{figure}
\end{minipage}
\end{tabular}
\end{frame}

@@

\begin{frame}
\frametitle{Mathematic foundation}
\framesubtitle{Algorithm for LBM with level set}
\begin{itemize}
\item<1-> Create initial interface
\item<2-> Run level set method to calculate surface description
\item<3-> Run LBM for a prescribed number of steps
\item<4-> Run level set method for the same time interval
\item<5-> Repeat until end of simulation
\end{itemize}
\end{frame}

\section{Results}
\begin{frame}
\frametitle{Results}
\framesubtitle{Setup}
Transient couette flow with immersed drop.
\begin{tabular}{l l}
\begin{minipage}{0.5\textwidth}

\begin{figure}[h!]
\includegraphics[width=7cm]{plots/vel0.png}
\end{figure}
\end{minipage}
&
\begin{minipage}{0.5\textwidth}
\begin{figure}[h!]
  \includegraphics[width=4cm]{plots/interface_t0.png}
\end{figure}
\end{minipage}
\end{tabular}
\end{frame}

\begin{frame}

\begin{tabular}{l l}
\begin{minipage}{0.5\textwidth}

\begin{figure}[h!]
\includegraphics[width=7cm]{plots/vel4.png}
\end{figure}
\end{minipage}
&
\begin{minipage}{0.5\textwidth}
\begin{figure}[h!]
  \includegraphics[width=4cm]{plots/interface_t4.png}
\end{figure}
\end{minipage}
\end{tabular}

@@

\begin{frame}
\begin{tabular}{l l}
\begin{minipage}{0.5\textwidth}
\begin{figure}[h!]
\includegraphics[width=7cm]{plots/vel10.png}
\end{figure}
\end{minipage}
&
\begin{minipage}{0.5\textwidth}
\begin{figure}[h!]
  \includegraphics[width=4cm]{plots/interface_t10.png}
\end{figure}
\end{minipage}
\end{tabular}
\dots more to come 
\end{frame}

\section{Outlook}
\begin{frame}
\frametitle{Outlook}
\textit{\dots so far intermediate results}

\vspace{.8cm} 
Outlook:
\begin{itemize}
\item<1-> Add correction term to prevent mass loss
\item<2-> Reduce computational effort: Store Level-Set function only in narrow band around interface, Adalsteinsson and Sethian
\item<3-> $\rightarrow$ \dots ?
%TODO: Quellen als Footnotes + Uebersichtsfolie
%\item<3-> Include thermal flow (simulate e.g. lava lamp) / Include gravity
\end{itemize}

\begin{frame}
\frametitle{References}
\begin{thebibliography}{etw}
   \bibitem{thoemmes}
     Thömmes, Guido, et al. "A lattice Boltzmann method for immiscible multiphase flow simulations using the level set method." Journal of Computational Physics 228.4 (2009): 1139-1156
     \bibitem{adalseth}
     D. Adalsteinsson, and J. A. Sethian. A fast level set method for propagating interfaces. J. Comput. Phys., 118(2):269-277, 1995
   \bibitem{mitchell}
     Mitchell, Ian, A Toolbox of Level Set Methods
\end{thebibliography}
\end{frame}
